---
title: "Barragem Doutor"
author: "Francisco d'Albertas"
date: "17/12/2021"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

#==== pacotes ==================================================================

library(sf)
library(dplyr)
library(raster)
library(flextable)
#===============================================================================

# caminho raiz

p <- "G:/Meu Drive/aquaflora"

# recortando raster

dir.create (file.path(p,"/temp"), showWarnings = FALSE)

#sets temp directory

rasterOptions(tmpdir=file.path(p,"/temp")) 

# knitr::knit("G:\\Meu Drive\\aquaflora\\aquaflora\\santana\\relatorio_santana.Rmd")
# rmarkdown::render("G:\\Meu Drive\\aquaflora\\aquaflora\\santana\\relatorio_santana.Rmd")
```

## Area da mancha de inundacao

8785.867719

## Area total dos municipios

Area total em ha baseada no shape
```{r echo=FALSE}

# esse eh pro IBGE

limites <- st_read(file.path(p,"aquaflora","doutor","municipios"))
 
limites$Relatorio <- "Doutor"

limites_ae <- limites %>% 
  group_by(Relatorio) %>%
  summarise()

area_total_mun <- as.numeric(st_area(limites_ae))/10^4

area_total_mun
```

## Populacao total!

## Usos do solo mapbiomas

- Cortei o mapa de usos do mapbiomas pro limite do relatorio Santana.


```{r include=FALSE}

limites_mapbio <- st_read(file.path(p,"aquaflora","doutor","limite"))


mapbio <- raster(file.path(p,"mapbiomas",
                         "mapbiomas-brazil-collection-60-minasgerais-2020.tif"))

#limites_ae <- st_zm(barragemlat,drop=T)
usos_c <- crop(x = mapbio,limites_mapbio)
usos_m <- mask(x = usos_c,limites_ae)



```

- Com o shape cortado, calculei os usos:

```{r include=FALSE}

# funcao pro calculo

empresa <- "Vale"
escala <- "bacia_modificada"

area_classes <- function(empresa,lista_rasters,escala){
  
  
    for(i in seq(1,length(lista_rasters),1)){
    
      r <-lista_rasters[[i]]# substituir por lista_rasters
      
      mini_rast <- r/r
      
      # get the total number of cells for this shape (save for below)
      
      total_cells <- cellStats(mini_rast, 'sum')
      
      # multiply by original - to make a mask layer from your original raster
      
      my_cutout <- r*mini_rast
      
      # Count the number of cells for each discrete landcover class
      
      in_this_poly_unit <- freq(my_cutout)
      
      #Divide the cell count for each of these classes by the total 
      # number of cells in your current shape
      
      class_percents <- in_this_poly_unit[,2]/total_cells
      
      
      # calculating area
      
      area <- as.data.frame(in_this_poly_unit) # exclui NAs!
      area <- area[!is.na(area$value),]
      area$area_ha <- area$count*0.09
      # make the function return a dataframe with
      # % landcover in each class for each shape
      output <- data.frame(classe_MapBiomas=in_this_poly_unit[,1], portion=class_percents)
      output <- output[!is.na(output$classe_MapBiomas),]
      output <- cbind(output,area[,3])
      names(output)[3] <- "area_ha"
      output$cluster <- names(r)
      #area=area
      codigo_mapbio <- data.frame(classe= c("Formacao Florestal"," Formacao Savanica","Silvicultura"," Campo Alagado","Formacao Campestre"," Pastagem"," Mosaico de Agricultura e Pastagem","Area Urbanizada","Outras Áreas nao Vegetadas","Afloramento Rochoso","Mineracao","Agua","Soja","Cafe"),classe_MapBiomas=output$classe_MapBiomas)
      output <- left_join(output,codigo_mapbio)
      return(output)
      write.csv(output,file.path(p,"aquaflora","doutor",
                                 paste0(empresa,"_tabela_usos_","Doutor",escala,".csv")))
  }

}


usos_Doutor <- area_classes(empresa = "Vale",lista_rasters = list(usos_m),escala = "bacia_modificada")

usos_Doutor <- usos_Doutor[,c(1:3,5)]
usos_Doutor$portion <- round(usos_Doutor$portion,2)

```


```{r echo=FALSE}

ft <- flextable(usos_Doutor)

ft


```

### Area total usando o mapbiomas

```{r echo=FALSE}
sum(usos_Doutor$area_ha)
```

### Soma agricultura,floresta,campos,pastagens,savana, silvicultura

```{r}
#sum(usos_Santana$area_ha[usos_Santana[c(1:3,5:7)])

sum(usos_Doutor[c(1:3,5:7),3])

```



## Usos agricolas do IBGE

```{r include=FALSE}

municipios <- limites$GEOCODIGO

lista_dados <- list.files(file.path(p,"tabelas_IBGE","bacias_7"),pattern = ".csv",full.names = T)

lista_dados_agg <- grep(lista_dados,pattern = "agg",value = T)

lista_mun <- lista_dados[!lista_dados %in% lista_dados_agg]


df_mun <- lapply(lista_mun,function(x)read.csv(x))


df <- do.call(rbind,df_mun)

df_sub <- df %>% filter(geocodigo %in% municipios)

# agregar por cultivo

names(df_sub)[2] <- "cultivo"

cultivos <- df_sub %>%
  group_by(cultivo)%>%
  filter(cultivo!="Café (em grão) Arábica")%>%
  summarise(area_total=sum(area_total))%>%
  filter(area_total!=0)


cultivos$prop <- round(cultivos$area_total/sum(cultivos$area_total),2)

empresa <- "Vale"
escala <- "bacia_modificada"

write.csv(cultivos,file.path(p,"aquaflora","doutor",                     paste0(empresa,"_culturas_IBGE_","doutor",escala,".csv")),
          row.names = F)





```

```{r echo=FALSE}

cultivos2 <- data.frame(cultivo="total",area_total=sum(cultivos$area_total),prop=1)

# reordenando

cultivos <- cultivos %>% arrange(desc(area_total))

cultivos3 <- rbind(cultivos,cultivos2)%>%
  arrange(desc(area_total))


ft2 <- flextable(cultivos3)

ft2

```

## Area agricola com menos de 2km distancia veg

```{r echo=FALSE}

eucdist <- raster(file.path(p,"aquaflora","doutor","eucdist","Doutor_dist_Veg.tif"))

m <- c(0,0,NA, 1,2000,1,2000,76440,NA)

rclmat <- matrix(m, ncol=3, byrow=TRUE)

r_rc <- reclassify(eucdist, rclmat)

sum_raster_cells <-cellStats(r_rc, 'sum')

sum_raster_cells*0.09

```



```{r include=FALSE}
#removes entire temp directory without affecting other running processes
unlink(file.path(file.path(p,"/temp")), recursive = TRUE)
```

